---
title: "EDA"
output: html_document
date: "2022-10-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose
Generate a bump plot of f1 fantasy rankings.

# Import / Setup
## Packages

```{r packages}
if (!require("pacman")) install.packages("pacman")

pacman::p_load(
    'dplyr',
    'forcats',
    'ggplot2',
    'janitor',
    'lubridate',
    'readr',
    'stringr'
)

```

## Data
```{r import_data}
url <- 'https://docs.google.com/spreadsheets/d/17xV0x__1V6_5GnWVG04m3g2es39yrqpZFnipXtswehc/edit#gid=0'
ssid <- googlesheets4::as_sheets_id(url)

points.df <- googlesheets4::range_speedread(
    ss = ssid, 
    range = "Sheet1!A:J", 
    col_types = readr::cols(
        .default      = readr::col_integer(), 
        Date          = readr::col_date('%m/%d/%Y'),
        Track         = readr::col_character(),
        Location      = readr::col_character(),
        'Mega Driver' = readr::col_character()
        )
)


points.df
```

# Processing
## Clean Column Names
```{r}
points.df <- points.df %>%
    janitor::clean_names()
```


## Count Race Number
```{r}
points.df <- points.df %>%
    arrange(date) %>% 
    mutate(
        race_number = row_number()
    )
```

## Long Form Dataset
```{r}
points.df_long <- points.df %>% 
    tidyr::pivot_longer(
        cols      = !c(location, date, track, race_number), 
        names_to  = 'player',
        values_to = 'points'
        )
```

## Create Name from Label
```{r}
points.df_long <- points.df_long %>%
  mutate(
      player_name = stringr::str_split_i(player, pattern = '_', 1)
    )
```

## Cummulative Points
```{r}
df_f1_cum <- points.df_long %>% 
    dplyr::arrange(date) %>% 
    dplyr::group_by(player) %>% 
    dplyr::mutate(
        points_cum = cumsum(points)
    )
```

## Point Differential
```{r}
df_f1_point_diff <- df_f1_cum %>% 
    group_by(race_number) %>% 
    mutate(
        max_points = max(points_cum),
        max_points_delta = points_cum - max_points
    )
```


## Rank
```{r}
df_f1_rank <- df_f1_cum %>% 
    dplyr::arrange(
        date, player, points_cum
    ) %>% 
    dplyr::group_by(
        date
    ) %>% 
    mutate(rank = rank(-points_cum, ties.method = 'min'))
```


# Visuals
## Visual Specifics
### Define Color Palette
```{r}
color_pal <- c(
    '#001219',
    '#005F73',
    '#0A9396',
    '#94D2BD',
    '#EE9B00',
    '#BB3E03',
    '#9B2226'
)

players <- c(
    'scott',
    'erin',
    'kevin', 
    'sydney',
    'evan',
    'kelsey',
    'margret'
)

# Assign names to colors
names(color_pal) <- players
```

### Defined Shared Inputs
```{r}
races <- df_f1_rank %>% 
    ungroup() %>%
    tidyr::drop_na() %>% 
    select(location) %>% 
    distinct() %>% 
    pull()
```


## Bump Chart
### Processing
#### Drop NA Rows
```{r}
df_plot <- df_f1_rank %>% 
    tidyr::drop_na()

df_plot
```

#### Create Dataset for Far Right Side Labels
```{r}
df_name <- df_plot %>% 
    filter(
        race_number == max(df_plot$race_number)
    ) %>% 
    mutate(
        label = str_to_title(player_name)
    )

df_name
```

### Generate Plot
```{r fig.height=10}
p <- ggplot(
    data    = df_plot, 
    mapping = aes(
        x     = race_number,
        y     = rank,
        color = player_name,
        )
    ) +
    geom_point(size = 5) +
    geom_line(size = 2.5) +
    geom_label(
        data = df_name,
        mapping = aes(
            x = race_number, 
            y = rank, 
            label = label
        ),
        hjust = 0, 
        nudge_x = 0.5, 
        size = 13, 
        label.size = NA
        ) +
    scale_x_continuous(
        breaks = 1:max(df_plot$race_number),
        labels = races, 
        ) +
    scale_y_reverse(breaks = 1:7) +
    scale_color_manual(values = color_pal) +
    # scale_color_brewer(
    #     type = 'qual',
    #     palette = 'Dark2'
    #     ) +
    labs(
        title = 'F1 Fantasy Ranking By Race', 
        x = '',
        y = ''
    ) +
    coord_cartesian(clip = 'off') +
    theme_classic() +
    theme(
        axis.line = element_line(size = 1.5),
        axis.text.x = element_text(
            hjust = 1,
            angle = 45, 
            size = 25
            ),
        axis.text.y = element_text(
            hjust = 0,
            size = 35,
        ),
        axis.ticks = element_blank(),
        legend.position="none", 
        plot.margin = margin(
            t = 10, r = 100, b = 10, l = 10, 
            unit = "pt"
            ),
        plot.title = element_text(
            hjust = 0,
            size = 40
            )
        )

p
```

### Save
```{r}
filename_prefix = 'f1_fantasy_rank'

# Define File Name
race_location <- df_plot %>% 
    tail(1) %>% 
    pull(location)

race_date <- df_plot %>% 
    tail(1) %>% 
    pull(date)
    
file_name_wo_ext <- paste(
    filename_prefix,
    race_date,
    race_location,
    sep = '_'
    )

file_name_wo_ext <- gsub(" ", "-", file_name_wo_ext)

file_name <- paste(
    file_name_wo_ext,
    '.png',
    sep = ''
)

# Save copy with race specific name
ggsave(
  filename = file_name,
  plot     = p,
  device   = 'png',
  path     = './../data/plots',
  scale = 1,
  width = 16.2,
  height = 10,
  units = "in",
  dpi = 300,
  limitsize = TRUE
)

# Save general copy
ggsave(
  filename = 'f1_fantasy_rank.png',
  plot     = p,
  device   = 'png',
  path     = './../data/plots',
  scale = 1,
  width = 16.2,
  height = 10,
  units = "in",
  dpi = 300,
  limitsize = TRUE
)
```

## Point Differential
### Processing
#### Drop NA Rows
```{r}
df_plot <- df_f1_point_diff %>% 
    tidyr::drop_na() %>% 
    mutate(
        max_points_delta_abs = abs(max_points_delta)
    )
```

#### Create Dataset for Far Right Side Labels
```{r}
df_name <- df_plot %>% 
    filter(
        race_number == max(df_plot$race_number)
    ) %>% 
    mutate(
        label = str_to_title(player_name)
    )
```

### Generate Plot
```{r  fig.height=10, fig.width= 16}
y_axis_bound <- max(df_plot$max_points_delta_abs)
y_axis_breaks <- seq(from = 100, to = y_axis_bound, by = 100)


p <- ggplot(
    data    = df_plot, 
    mapping = aes(
        x     = race_number,
        y     = max_points_delta_abs,
        color = player_name,
        )
    ) +
    geom_point(size = 5) +
    geom_line(size = 2.5, ) +
    ggrepel::geom_text_repel(
        data = df_name,
        mapping = aes(
            x = race_number, 
            y = max_points_delta_abs, 
            label = label
        ),
        hjust = 0, 
        nudge_x = 01,
        size = 13, 
        segment.color = NA,
        direction = "y"
        ) +
    scale_x_continuous(
        breaks = 1:max(df_plot$race_number),
        labels = races, 
        limits = c(1, length(races) + 4)
        ) +
    scale_y_reverse(
        breaks = y_axis_breaks,
        # labels = y_axis_breaks
    ) +
    scale_color_manual(values = color_pal) +
    labs(
        title = 'F1 Fantasy Point Differential To Leader', 
        x = '',
        y = 'Points Behind Leader'
    ) +
    theme_classic() +
    theme(
        axis.line = element_line(size = 1.5),
        axis.title.y = element_text(
            hjust = 0.95,
            size = 35
            ),
        axis.text.x = element_text(
            hjust = 1,
            angle = 45, 
            size = 25
            ),
        axis.text.y = element_text(
            hjust = 1,
            size = 35,
        ),
        axis.ticks = element_blank(),
        legend.position="none", 
        plot.margin = margin(
            t = 10, r = 10, b = 10, l = 10, 
            unit = "pt"
            ),
        plot.title = element_text(
            hjust = 0,
            size = 40
            )
        )

p
```

### Save
```{r}
filename_prefix = 'f1_fantasy_point_diff'

# Define File Name
race_location <- df_plot %>% 
    tail(1) %>% 
    pull(location)

race_date <- df_plot %>% 
    tail(1) %>% 
    pull(date)
    
file_name_wo_ext <- paste(
    filename_prefix,
    race_date,
    race_location,
    sep = '_'
    )

file_name_wo_ext <- gsub(" ", "-", file_name_wo_ext)

file_name <- paste(
    file_name_wo_ext,
    '.png',
    sep = ''
)

# Save copy with race specific name
ggsave(
  filename = file_name,
  plot     = p,
  device   = 'png',
  path     = './../data/plots',
  scale = 1,
  width = 16.2,
  height = 10,
  units = "in",
  dpi = 300,
  limitsize = TRUE
)

# Save general copy
ggsave(
  filename = 'f1_fantasy_point_diff.png',
  plot     = p,
  device   = 'png',
  path     = './../data/plots',
  scale = 1,
  width = 16.2,
  height = 10,
  units = "in",
  dpi = 300,
  limitsize = TRUE
)
```

